---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r,echo=FALSE,results='hide',message=FALSE}
library(dplyr)
library(gridExtra)
library(ggplot2)
```

## Loading and preprocessing the data
Data loading and converting `date` variable to date format.
```{r}
data <- read.csv('activity.csv',col.names = c('steps','date','interval'),
                 na.strings = 'NA')
data <- tbl_df(data)
data <- mutate(data, date = as.Date(date,'%Y-%m-%d'))
data
```

## What is mean total number of steps taken per day?
Grouping by `date` and getting the total of steps/day.
```{r}
steps_day <- data %>% 
    group_by(date) %>% 
    summarise(steps = sum(steps,na.rm = TRUE))
```
Making the histogram:
```{r,message=FALSE}
plt1 <- ggplot(steps_day,aes(x = steps)) + 
    geom_histogram() +
    labs(x = 'Steps',y = 'Counts')
print(plt1)
```

With **mean** `r round(mean(steps_day$steps),2)` and **median** 
`r median(steps_day$steps)` steps.

## What is the average daily activity pattern?
Grouping by `interval` and getting the mean of steps/day.
```{r}
steps_interval <- data %>% 
    group_by(interval) %>% 
    summarise(steps = mean(steps,na.rm = TRUE))
```
Plotting:
```{r,message=FALSE}
plt2 <- ggplot(steps_interval,aes(x = interval,y = steps)) + 
    geom_line() +
    labs(x = 'Interval',y = 'Steps')
print(plt2)
```

And the 5-minute interval with the **maximum average** of steps is 
`r steps_interval$interval[which.max(steps_interval$steps)]`.

## Imputing missing values
The total number of *NA*'s is `r sum(!complete.cases(data))`, filling those 
values with the step's mean for that interval. Creating a new dataset with the *NA*'s
values filled in.
```{r}
data
nas = which(is.na(data$steps) == TRUE)
n = length(nas)
for (i in 1:n) {
    j = nas[i]
    data$steps[j] <- steps_interval$steps[which(steps_interval$interval == data$interval[j])]
}
```
Double checking *NA*'s values: `r sum(!complete.cases(data))`.  
Plotting the histogram:
```{r,message=FALSE}
steps_day <- data %>% 
    group_by(date) %>% 
    summarise(steps = sum(steps))
plt3 <- ggplot(steps_day,aes(x = steps)) + 
    geom_histogram() +
    labs(title= 'NA filled',x = 'Steps',y = 'Counts')
plt1 <-  plt1 + ggtitle('NA ignored')
grid.arrange(plt1, plt3, nrow = 1)
```

With **mean** `r format(round(mean(steps_day$steps),2),scientific=F)` and **median** 
`r format(median(steps_day$steps),scientific=F)` steps.  
The effect of imputing data is a change in the mean and median of the total steps per day.  

## Are there differences in activity patterns between weekdays and weekends?
Making weekdays factor `weekday` or `weekend`:
```{r}
data <- mutate(data, day =  tolower(weekdays(data$date)))
data$day <- as.factor(ifelse(data$day=="saturday"|data$day=="sunday","weekend","weekday"))
patterns <- summarise(group_by(data,interval,day),steps = mean(steps))
plt4 <- ggplot(patterns,aes(x=interval,y=steps)) +
    geom_line() +
    facet_grid(day~.)
print(plt4)
```
